---
title: "OCR+PER-Pipeline-FOUR-INTERVALS"
output: 
  html_document:
    number_sections: true
params:
  input_folder: "4_intervals"
editor_options: 
  chunk_output_type: console
---
<br>
<br>
<br>
<br>

# HOW TO USE

Step 1: Make a new folder in the Data/INPUT folder and place your excel files there <br>
Step 2: Make sure that you have not opened any of the files in the folder <br>
Step 3: Write the name of the folder above. Use quotes like this "My_folder" <br>
Step 4: Run the code chunk below <br>



```{r eval = FALSE, include=FALSE}
# Set path and load the package "here"

here::i_am("Experiments/OCR+PER-Pipeline.Rmd")

if(!("here" %in% installed.packages()[, "Package"])) {
  install.packages("here")
}
library(here)       # Used to set the path to the package dir. on the machine 



# Check if all packages are installed
source(here("R", "src", "config_file.R"))


# Load libraries
library(tools)
library(reshape2)
library(reshape)
library(tidyverse)
library(readxl)     # Read xlsx file
library(knitr)      # Used for tables 
library(kableExtra)
library(shiny)



# Get input folder
input_folder <- params$input_folder

# Create output folder

output_folder <- sprintf("%s_OUTPUT", input_folder)
o_path <- here("Data", "OUTPUT", output_folder)

if(!dir.exists(o_path)){
  dir.create(path = o_path)
}


# source the functions used in analysis
source(here("R", "src", "analysis_source_functions_FOUR_INTERVALS.R"))

# INPUT_PATH  <- (here('/Data/INPUT/ConBIS')) # path to the folder of interest
INPUT_PATH  <- (here('Data', "INPUT", input_folder)) # path to the folder of interest
# OUTPUT_PATH <- (here('/Data/OUTPUT/HF')) 
OUTPUT_PATH <- (here('Data', "OUTPUT", output_folder))


# Delete old knit files
knit_cache <- here('Experiments', "OCR+PER-Pipeline_cache")
knit_files <- here("Data", "OUTPUT", output_folder, "OCR+PER-Pipeline_files")
knit_folders <- c(knit_cache, knit_files)
for (folder in knit_folders) {
  if (file.exists(folder))
  unlink(folder, recursive = T)
}


# Initialize the knitting process
rmarkdown::render(here('Experiments', "OCR+PER-Pipeline-FOUR-INTERVALS.Rmd"), output_dir = OUTPUT_PATH)
```


```{r, message=FALSE, warning=FALSE, echo=FALSE, paged.print=FALSE}
# Read The the separate xlsx files from folder, Specify entire path to folder. 
source(here("R", "src", "analysis_source_functions_FOUR_INTERVALS.R"))
data <- read_xlsx_set(INPUT_PATH, ".xlsx")  
dm <- data$rates

```

# Remove wells based on pressure measurements + correct background

```{r echo=FALSE}
data$Hg_list %>%  
  kable(caption="Wells removed. Out od 140-160 mmHg interval") %>% 
  kable_styling()

```

```{r echo=FALSE}
data$OCR_background %>%  
  kable(caption="Background measurements not included in OCR correction") %>% 
  kable_styling()

```

```{r echo=FALSE}
data$PER_background %>%  
  kable(caption="Background measurements not included in PER correction") %>% 
  kable_styling()

```

```{r echo=FALSE}
data$Zero_measurements %>%  
  kable(caption="Empty measurements") %>% 
  kable_styling()

```

<br>
<br>

# OCR 
## Filter out irrelevant samples 
```{r  echo=FALSE}
# perep for OCR interval 5 not used  
d_OCR  <- dm
# filter out samples with any of intervals having less than 8 measurements 
sumar <- d_OCR %>%
  group_by(sample_id, Interval) %>% 
  summarise(n = n()) %>% 
  filter(n < 18) # les then 18 so it corespods to at least 6 different wells
# number of intervals has to be 4 or discarted 
inter <- d_OCR %>% 
  group_by(sample_id) %>% 
  summarise(n_inter = length(unique(Interval))) %>% 
  filter(n_inter != 4)

d_OCR <- d_OCR %>% 
  filter(! sample_id %in% unique(sumar$sample_id))

d_OCR <- d_OCR %>% 
  filter(! sample_id %in% unique(inter$sample_id))

paste0("Samples removed due lack of usable measurements(less than 8  in any interval)")
unique(sumar$sample_id)
paste0("Samples removed because there are no 4 intervals )")
unique(inter$sample_id)
```

## Identify Outliers 
Removed outliers in each iteration:
```{r echo=FALSE, echo=FALSE}
# Identify Outliars 
# DT = loaded dataframe 
# cut.well = threshold for well outliaer, 
# cut.point = threshold for point outliaer
# x = Variable: "LOCR", "OCR

dr_ocr <- idfy_sinleP_outlier(d_OCR, cut.point = 6, x = "LOCR") 
# clean enviroment 
rm(d_OCR)
# write the data into file 
write_csv(dr_ocr, paste0(OUTPUT_PATH,"/Data_removed_OCR.csv"))  
```

```{r echo=FALSE, warning=FALSE, echo=FALSE}
# amount of removed outliars
dr_ocr %>% 
  group_by(sample_id, is.out.p) %>% 
  summarise(n = n()) %>% 
  group_by(sample_id) %>% 
  mutate(size = sum(n), 
         ptg = n / size *100,
         percentage = paste(format(ptg, digits= 1, nsmall=2),"%")) %>% 
  filter(is.out.p == T) %>% 
  arrange(desc(ptg)) %>% 
  select(sample_id, percentage) %>%
  kable(caption="Amount of removed outliers across samples") %>% 
  kable_styling(full_width = F)


```

## Print schemes of removed points for all samples 
```{r echo=FALSE, echo=FALSE, cache = TRUE}

for (smpl in unique(dr_ocr$sample_id)) {
  d    <- filter(dr_ocr, sample_id == smpl)
  ptg  <- nrow(filter(d, is.out.p == T))/nrow(d)*100
  plot <- ggplot(d, aes(Time, OCR))+
            ggtitle(paste0("ID: ", smpl, " |  Outliers: ", format(ptg, digits = 3), "%"))+
            geom_line(aes(group = Well),size = 0.2, color = "grey") +
            geom_point(aes(Time, OCR, color = is.out.p))+
            guides(color=guide_legend(title="Outlier"))+
            xlab("Measurement")+
            ylim(0,15)+
            ylab("OCR (Normalized)")

  print(plot +theme_bw())
}
 # Violin plot LOCR
#  for (smpl in unique(dr_ocr$sample_id)) {
#    d    <- filter(dr_ocr, sample_id == smpl)
#    ptg  <- nrow(filter(d, is.out.p == T))/nrow(d)*100
#    plot <- ggplot(d, aes(Time, LOCR))+
#              ggtitle(paste0("ID: ", smpl, "  out: ", format(ptg, digits = 3), "%"))+
#              geom_violin(aes(Interval, LOCR))+
#              geom_point(aes(Interval, LOCR, color = is.out.p))+
#              xlab("Interval")#+
#  
#    print(plot +theme_bw())
# }
```


## Bioenergetics and Estimates 
### Estimates
Estimates are taken by median of measurements from each interval in sample. Estimates are  stored together with other details about the intervals in files called <Method>-Estimated_values.csv in OUTPUT folder.

If you want to see or change how they are computed see the function compute_bioenergetics_ ()in file analysis_source_functions.R

```{r echo=FALSE}
norm.be <- compute_bioenergetics_(dr_ocr, "OCR")
log.be  <- compute_bioenergetics_(dr_ocr, "LOCR") 

write_csv(norm.be$estimates, paste0(OUTPUT_PATH,"/OCR-Estimated_values.csv"))
write_csv(log.be$estimates, paste0(OUTPUT_PATH,"/LOCR-Estimates_values.csv"))
write_csv(norm.be$bioenergetics, paste0(OUTPUT_PATH,"/OCR-BE-table.csv"))
write_csv(log.be$bioenergetics, paste0(OUTPUT_PATH,"/LOCR-BE-table.csv"))

```

```{r echo=FALSE, echo=FALSE}
# Print interval estimates 
norm.be$estimates[1:4,] %>%
  kable(digits = 3, caption = "Example of estimates file ") %>%
  kable_styling(full_width = F)
```


```{r echo=FALSE}
# norm estimates boxplots
ggplot(norm.be$estimates)+
  ggtitle("OCR Estimates natural scale")+
  geom_boxplot(aes(Interval,mean), width = 0.6, outlier.size = -1, show.legend = FALSE, fill = "pink")+
  geom_jitter(aes(Interval,mean), width = 0.2, show.legend = FALSE, size = 0.5 )+
  xlab("Intervals")+
  ylab("OCR")+
  theme_bw()

# log estimates boxplots
ggplot(log.be$estimates)+
  ggtitle("logOCR Estimates")+
  geom_boxplot(aes(Interval,mean), width = 0.6, outlier.size = -1, show.legend = FALSE, fill = "pink")+
  geom_jitter(aes(Interval,mean), width = 0.2, show.legend = FALSE, size = 0.5 )+
  xlab("Intervals")+
  ylab("logOCR")+
  theme_bw()

```


### Bioenergetics
Bioenergetics are computed from estimates, there there are two sets of bioenergetics provided. 

  1. Ratio Based, computed from log(OCR) estimates Located in OUTPUT/LogOCR-BioEnergetics.csv
  
  2. Difference based, computed from OCR esimates Located in OUTPUT/OCR-BioEnergetics.csv
If you want to see or change how they are computed see the function compute_bioenergetics_() in file analysis_source_functions.R

```{r message=FALSE, warning=FALSE, echo=FALSE, fig.width=10}
# ADD coefficients of variation
# normal scale Bioenergetics 

n.bio <- melt(norm.be$bioenergetics)

n.bio %>% 
  filter(variable != "Other") %>% 
  ggplot(aes(variable, value, fill = variable ))+
  ggtitle("Difference based natural scale Bio-Energetics OCR")+
  geom_boxplot(width = 0.5, outlier.size = -1, alpha = 0.7)+
  geom_jitter(width = 0.1, show.legend = FALSE, size = 0.5 )+
  xlab("Bio-Energetics")+
  ylab(" ")+
  theme_bw()


# log scale Bioenergetics 
l.bio   <- melt(log.be$bioenergetics )
l.bio %>% 
  filter(variable != "Other") %>% 
  ggplot(aes(variable, value, fill = variable))+
  ggtitle("Ratio based log scale Bio-Energetics OCR (folds)")+
  geom_boxplot( width = 0.5, outlier.size = -1, alpha = 0.7)+
  geom_jitter( width = 0.1, show.legend = FALSE, size = 0.5 )+
  xlab(" Log Bio-Energetics")+
  ylab(" ")+
  theme_bw()


  

```

```{r echo=FALSE}
# write difference based bioenergetics
# uncomment and export be.er if you want to get standard errors of bioenergetics 
# st.err <- melt(norm.be$standard.errors, value.name = "SE", variable.name = "Bio.e" )
# be.er  <- left_join(n.bio, st.err, by = c("Sample", "variable"))
# be.er  <- arrange(be.er, Sample)
write_csv(n.bio, paste0(OUTPUT_PATH,"/OCR-BioEnergetics.csv"))

# write ratio based bienergetics
# # uncomment and export be.er if you want to get standard errors of bioenergetics 
# st.err <- melt(log.be$standard.errors, value.name = "SE", variable.name = "variable" )
# be.er  <- left_join(l.bio, st.err, by = c("Sample", "variable"))  
# be.er <- arrange(be.er, Sample)
write_csv(l.bio, paste0(OUTPUT_PATH,"/LogOCR-BioEnergetics.csv"))
  

```

```{r echo=FALSE}
# Print interval estimates 
l.bio[1:4,] %>%
  kable(digits = 3, caption = "Example of Bioenergetics file ") %>%
  kable_styling(full_width = F)
```

### Groups 

If the sample_id of your samples contains letters A or B following graph will plot bioenergetics of the two grups. 
It is suggested to use this plot only when you are looking at grouped samples and project name shouldn't contain letters A or B. Any sample that is not labeled will be part of B group. 

You can change the A and B letters in the markdown code to compare other groups. 

```{r, echo=FALSE, fig.width = 10 }

# difference based BE
n.bio %>% 
  mutate(Group = ifelse(grepl("A", Sample), "A", "B")) %>% # change "A" and "B" for any character or string
  ggplot(aes(Group, value, fill = variable))+
  ggtitle("OCR Normal scale Bio-Energetics biological groups ")+
  geom_boxplot(width = 0.5, outlier.size = -1, alpha = 0.7)+
  geom_jitter(width = 0.1, show.legend = FALSE, size = 0.5)+
  xlab("Bio-Energetics")+
  ylab("OCR")+
  facet_grid(. ~ variable ) +
  theme_bw()

# ratio based BE
l.bio %>% 
  mutate(Group = ifelse(grepl("A", Sample), "A", "B")) %>% # change "A" and "B" for any character or string
  ggplot(aes(Group, value, fill = variable))+
  ggtitle("OCR Log scale Bio-Energetics biological groups ")+
  geom_boxplot(width = 0.5, outlier.size = -1, alpha = 0.7)+
  geom_jitter(width = 0.1, show.legend = FALSE, size = 0.5)+
  xlab("Bio-Energetics")+
  ylab("OCR")+
  facet_grid(. ~ variable ) +
  theme_bw()


```


# PER

## Filter out irrelevant samples 

```{r echo=FALSE}
# perep for PER interval 1 not used  
d_PER <- dm %>% filter(!Interval %in% c("Int3", "Int4"))
                        
# filter out samples with any of intervals having less than 8 measurements 
sumar <- d_PER %>%
  group_by(sample_id, Interval) %>% 
  summarise(n = n()) %>% 
  filter(n < 8 | Interval == "Int1" & n < 18 | Interval == "Int2" & n < 18)

# number of intervals has to be 2 or discarted 
inter <- d_PER %>% 
  group_by(sample_id) %>% 
  summarise(n_inter = length(unique(Interval))) %>% 
  filter(n_inter != 2)

d_PER <- d_PER %>% 
  filter(! sample_id %in% unique(sumar$sample_id)) %>% 
  filter(! sample_id %in% unique(inter$sample_id))

paste0("Samples removed due lack of usable measurements(less than 8  in any interval)")
unique(sumar$sample_id)
paste0("Samples removed because it does not have 2 Intervals")
unique(inter$sample_id)
```

## Remove outliers
```{r echo=FALSE, echo=FALSE}
# Identify Outliars 
# DT = loaded dataframe 
# cut.well = threshold for well outliaer, 
# cut.point = threshold for point outliaer
# x = Variable: "LOCR" or "OCR" or "PER"

dr_per <- idfy_sinleP_outlier(DT = d_PER, cut.point = 5, x = "PER" )  # could Print arguments
# write the data into file 
write_csv(dr_per, paste0(OUTPUT_PATH,"/Data_removed_PER.csv"))  
```

```{r echo=FALSE, warning=FALSE, echo=FALSE}
#TO DO: convert graph to aditive add legends 
#TO DO: convert graph to aditive add legends 
dr_per %>% 
  group_by(sample_id, is.out.p) %>% 
  summarise(n = n()) %>% 
  group_by(sample_id) %>% 
  mutate(size = sum(n), 
         ptg = n / size *100,
         percentage = paste(format(ptg, digits= 1, nsmall=2),"%")) %>% 
  filter(is.out.p == T) %>% 
  arrange(desc(ptg)) %>% 
  select(sample_id, percentage) %>%
  kable(caption="Amount of removed outliers across samples") %>% 
  kable_styling(full_width = F)

```


## Print schemes of removed points for all samples
```{r echo=FALSE}
for (smpl in unique(dr_per$sample_id)) {
  plot <- ggplot(filter(dr_per, sample_id == smpl), aes(Measurement, PER))+
            ggtitle(paste0("SAMPLE:  ", smpl))+
            geom_line(aes(group = Well),size = 0.2, color = "grey") +
            geom_point(aes(Measurement, PER, color = is.out.p))+
            guides(color=guide_legend(title="Outlier"))+
            xlab("Measurement")
  
  print(plot +theme_bw())
}
```

## Bioenergetics and Estimates 

```{r echo=FALSE}
norm.be.PER <- compute_bioenergetics_(dr_per, "PER") 
log.be.PER  <- compute_bioenergetics_(dr_per, "LPER") 

write_csv(norm.be.PER$estimates, paste0(OUTPUT_PATH,"/PER-Estimated_values.csv"))
write_csv(log.be.PER$estimates, paste0(OUTPUT_PATH,"/LPER-Estimated_values.csv"))
write_csv(norm.be.PER$bioenergetics, paste0(OUTPUT_PATH,"/PER-BE-table.csv"))
write_csv(log.be.PER$bioenergetics, paste0(OUTPUT_PATH,"/LPER-BE-table.csv"))
```
### Estimates

```{r echo=FALSE}

# norm estimates boxplots
norm.be.PER$estimates %>% 
  ggplot(aes(Interval,mean))+
  ggtitle("PER Estimates natural scale")+
  geom_boxplot( width = 0.6, outlier.size = -1, show.legend = FALSE, fill="pink", alpha = 0.7)+
  # geom_line(aes(group = sample_id, Interval,mean), col = "grey", size = .1, show.legend = FALSE)+
  geom_jitter( width = 0.2, show.legend = FALSE, size = 0.5 , alpha = 0.7)+
  xlab("Intervals")+
  ylab("logPER")+
  theme_bw()

# log estimates boxplots
log.be.PER$estimates %>%   
  ggplot(aes(Interval,mean))+
  ggtitle("Log PER Estimates")+
  geom_boxplot(width = 0.6, outlier.size = -1, show.legend = FALSE, fill="pink", alpha = 0.7)+
  # geom_line(aes(group = sample_id, Interval,mean), col = "grey", size = .1, show.legend = FALSE)+
  geom_jitter(width = 0.2, show.legend = FALSE, size = 0.5 )+
  xlab("Intervals")+
  ylab("logPER")+
  theme_bw() 
  
  
  
  
```


### Bioenergetics

```{r message=FALSE, warning=FALSE, echo=FALSE, fig.width = 8}
# ADD coefficients of variation
# normal scale Bioenergetics 

n.bio.PER <- melt(norm.be.PER$bioenergetics)
n.bio.PER %>% 
  filter(variable != "Other") %>% 
  ggplot(aes(variable, value, fill = variable ))+
  ggtitle("Difference based natural scale Bio-Energetics PER")+
  geom_boxplot(width = 0.5, outlier.size = -1, alpha = 0.7)+
  geom_jitter(width = 0.1, show.legend = FALSE, size = 0.5 )+
  xlab("Bio-Energetics")+
  ylab(" ")+
  theme_bw()


# log scale Bioenergetics 
l.bio.PER   <- melt(log.be.PER$bioenergetics )
l.bio.PER %>% 
  filter(variable != "Other") %>% 
  ggplot(aes(variable, value, fill = variable ))+
  ggtitle("Difference based log scale Bio-Energetics PER")+
  geom_boxplot(width = 0.5, outlier.size = -1, alpha = 0.7)+
  geom_jitter(width = 0.1, show.legend = FALSE, size = 0.5 )+
  xlab("Bio-Energetics")+
  ylab(" ")+
  theme_bw()



# add standard errors
# st.err <- melt(norm.be.PER$standard.errors, value.name = "SE", variable.name = "Bio.e" )
# be.er  <- left_join(n.bio.PER, st.err, by = c("Sample", "variable"))
# be.er  <- arrange(be.er, Sample)
write_csv(n.bio.PER, paste0(OUTPUT_PATH,"/PER-BioEnergetics.csv"))
# add standard errors
# st.err <- melt(log.be$standard.errors, value.name = "SE", variable.name = "variable" )
# be.er  <- left_join(l.bio.PER, st.err, by = c("Sample", "variable"))  
# be.er <- arrange(be.er, Sample)
write_csv(l.bio.PER, paste0(OUTPUT_PATH,"/logPER-BioEnergetics.csv"))




```

### Groups

```{r echo=FALSE, fig.width=8}
# difference based BE
n.bio.PER %>% 
  mutate(Group = ifelse(grepl("A", Sample), "A", "B")) %>% 
  ggplot(aes(Group, value, fill = variable))+
  ggtitle("PER Normal scale Bio-Energetics biological groups ")+
  geom_boxplot(width = 0.5, outlier.size = -1, alpha = 0.7)+
  geom_jitter(width = 0.1, show.legend = FALSE, size = 0.5)+
  xlab("Bio-Energetics")+
  ylab("PER")+
  facet_grid(. ~ variable ) +
  theme_bw()

# ratio based BE
l.bio.PER %>% 
  mutate(Group = ifelse(grepl("A", Sample), "A", "B")) %>% 
  ggplot(aes(Group, value, fill = variable))+
  ggtitle("PER Log scale Bio-Energetics biological groups ")+
  geom_boxplot(width = 0.5, outlier.size = -1, alpha = 0.7)+
  geom_jitter(width = 0.1, show.legend = FALSE, size = 0.5)+
  xlab("Bio-Energetics")+
  ylab("logPER")+
  facet_grid(. ~ variable ) +
  theme_bw()


```


# PER vs OCR

```{r echo=FALSE}
basal.OCR <- n.bio %>% 
  filter(variable == "Basal.Resp") %>% 
  mutate(Group = ifelse(grepl("A", Sample), "A", "B"))
basal.PER <- n.bio.PER %>% 
  filter(variable == "Basal.PER") %>% 
  mutate(Group = ifelse(grepl("A", Sample), "A", "B"))

total <- merge(basal.PER, basal.OCR, by = "Sample", suffixes = c(".PER", ".OCR"))

# total <- total %>% 
#   mutate(Group = ifelse(grepl("B", Sample), "Post", "Pre"))

ggplot(total)+
  ggtitle("Basal.PER vs Basal.OCR")+
  geom_point(aes(value.PER, value.OCR))+
  ylab("Basal OCR")+
  xlab("Basal PER")+
  theme_bw()
  
  
```

```{r echo=FALSE}
max.OCR   <- n.bio %>% filter(variable == "Maximal.Resp") %>% 
  mutate(Group = ifelse(grepl("A", Sample), "A", "B"))
max.PER <- n.bio.PER %>% filter(variable == "Max.PER") %>% 
  mutate(Group = ifelse(grepl("A", Sample), "A", "B"))

total_max <- merge(max.PER, max.OCR, by = "Sample", suffixes = c(".PER", ".OCR"))

# total <- total %>% 
#   mutate(Group = ifelse(grepl("B", Sample), "Post", "Pre"))

ggplot(total_max)+
  ggtitle("Max PER vs Max OCR")+
  geom_point(aes(value.PER, value.OCR))+
  ylab("Max OCR")+
  xlab("Max PER")+
  theme_bw()
  
  
```

# Stines test area. Must be deleted when pipeline is finished
<!-- # ```{r} -->
<!-- # E %>% -->
<!-- #   ggplot(aes(Interval,PER))+ -->
<!-- #   ggtitle("PER Estimates natural scale")+ -->
<!-- #   geom_boxplot( width = 0.6, outlier.size = -1, show.legend = FALSE, fill="pink", alpha = 0.7)+ -->
<!-- #   # geom_line(aes(group = sample_id, Interval,mean), col = "grey", size = .1, show.legend = FALSE)+ -->
<!-- #   geom_jitter( width = 0.2, show.legend = FALSE, size = 0.5 , alpha = 0.7)+ -->
<!-- #   xlab("Intervals")+ -->
<!-- #   ylab("logPER")+ -->
<!-- #   theme_bw() -->
<!-- # ``` -->

